---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Assignment 5: Analyzing repeated measures experiments {#assignment5}

## Assignment overview

In this assignment you will analyse and report on trial investigating the effect of resistance training volume on lean mass and muscle strength. The data are part of the `exscidata` package and can be accessed as `data("strengthvolume")` and `data("dxadata")`. Read the [instructions carefully](https://dhammarstrom.github.io/quant-methods-workshops/assignment-5.html)!


> Below you will find a basic outline of the report and example code that we worked on in class.

## Introduction

## Methods

### Participants and study overview

### Muscle strength and hypertrophy

### Data analysis and statistics

## Results

```{r}
#| label: "Load packages and data"
#| warning: false
#| message: false
#| echo: false
#| output: false



library(tidyverse); library(exscidata)
library(lme4)


## Import data set, extract leg lean mass per leg and remove observations
# that are include == "excl".

leanmass <- dxadata %>%
  select(participant:include, lean.left_leg, lean.right_leg) %>%
  pivot_longer(names_to = "leg", 
               values_to = "lean.mass", 
               cols = lean.left_leg:lean.right_leg) %>%
  mutate(leg = if_else(leg == "lean.left_leg", "L", "R"), 
         sets = if_else(multiple == leg, "multiple", "single")) %>%
  select(participant, time, sex, include, sets, leg, lean.mass) %>%
  filter(include == "incl") %>%
  
  ## Change levels for time and sets
  mutate(time = factor(time, levels = c("pre", "post")), 
         sets = factor(sets, levels = c("single", "multiple")), 
         leg = paste0(participant, leg)) %>%
  
  print()


# Create a wide data set of change scores per volume condition
# suitable for a t-test
leanmass_change_wide <- leanmass %>%
  pivot_wider(names_from = time, values_from = lean.mass) %>%
  mutate(change = post - pre) %>%
  
  select(participant, sex, sets, change) %>%
  pivot_wider(names_from = sets, 
              values_from = change) %>%
  
  mutate(diff = multiple - single) %>%
  
  print()


leanmass_change_wide_time <- leanmass %>%
  pivot_wider(names_from = time, values_from = lean.mass) %>%

  select(participant, sex, sets, pre, post) %>%
  
  print()




## Do a t-test on change scores per condition. 

## Alternatives (paired, equal variation)
tt1 <- with(leanmass_change_wide, t.test(multiple, single, paired = TRUE))

## The same model but with lm
m1 <- lm(diff ~ 1, data = leanmass_change_wide)

summary(m1)




### Mixed effects model
# A mixed effects model allowing for a random slope for time is very similar 
# to the t-test used above. 

m2 <- lmer(lean.mass ~ time * sets +( 1 + time |participant), 
           data = leanmass)

# Check the summary 
summary(m2)
# Check assumptions of equal residual variance
plot(m2)

## Mixed-Effects ANCOVA

# The lm model does not contain information on multiple
# observations per participant
m3 <- lm(post ~ pre + sets, data = leanmass_change_wide_time)


# A more correct model also include random intercepts per 
# participant. Mixed effects ANCOVA mnodel:
m4 <- lmer(post ~ pre + sets + (1|participant), 
           data = leanmass_change_wide_time)


summary(m3)

summary(m4)



### How to extract data from a model:

# m1 is a linear model of difference in differences
# we can use it for inference, extract average diff and confidence 
# intervals and p-value

# Save all elements of inline results
m1pval <- round(coef(summary(m1))[1, 4], 3)
m1est <- round(coef(summary(m1))[1, 1], 1)

m1ciu <- round(confint(m1)[1,2], 1)
m1cil <- round(confint(m1)[1,1], 1)

# Combine into an object
m1results <- paste0(m1est, 
                    ", 95% CI: [",
                    m1cil, 
                    ", ",
                    m1ciu, 
                    "], p = ",
                    m1pval
                    )





```

The average difference in lean mass changes between sets were `r m1results`.



```{r}

## Time points in strength data set

strengthvolume %>%
  distinct(exercise)



## Exploratory plot of strength data 

str <- strengthvolume %>%
  filter(include == "incl") %>%
  mutate(time = factor(time, levels = c("pre", "session1", 
                                        "week2", "week5", 
                                        "week9", "post"))) %>%
  print()


str %>%
  ggplot(aes(time, 
             load, 
             group = paste(participant, sets), 
             color = sets)) + 
  geom_line() + 
  facet_wrap(~ exercise, scales = "free")
  

## How many measurements per participant

str %>%
  filter(!is.na(load)) %>%
  group_by(participant, exercise, sets) %>%
  summarise(n = n() ) %>%
  ggplot(aes(n, participant, color = sets)) +
  geom_point() + 
  facet_wrap(~ exercise) + 
  theme(axis.text.y = element_blank())



## Use pre and post data 
# Combine pre data prior to data analysis
# per exercise, leg, participant, and sets

str %>%
  mutate(time = if_else(time %in% c("pre", "session1"), "pre", time)) %>%
  
  filter(time %in% c("pre", "post")) %>%
  
  summarise(load = max(load, na.rm = TRUE), 
            .by = c(participant, 
                    sex, 
                    time, 
                    sets,
                    exercise,
                    leg)) %>%
  
  print()











```





## Discussion

## Conclusion






